Red Wine Quality EDA by Anas Haneya
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(tidyr)
library(gridExtra)
library(dplyr)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
red <- read.csv('wineQualityReds.csv')
```

**Public Data Set about Red Wine with 13 variable and 1599 observation**

# Univariate Plots Section

```{r echo=FALSE, structure}
str(red)
```


```{r echo=FALSE, dim}
dim(red)
```


```{r echo=FALSE, summary}
summary(red)
```

```{r echo=FALSE, nulls_check}
sum(is.na(red))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, quality_plot}
ggplot(data = red, aes(x = as.integer(as.vector(quality)))) +
  geom_histogram() +
  ggtitle('Histogram of wine qualities distrubition')
```
** Most of the wine samples are centered on 5, 6, 7 the extreme low and extreme <br> low can be considered as outliers **

```{r echo=FALSE, message=FALSE, warning=FALSE, density}
ggplot(data = red, aes(x = density)) +
  geom_histogram() +
  ggtitle('Histogram of wine densities distrubition')
```
** Density of the wine is normally distributed **
```{r echo=FALSE, message=FALSE, warning=FALSE, fixed.acidity}
ggplot(data = red, aes(x = fixed.acidity)) +
  geom_histogram() +
  ggtitle('Histogram of wine fixed acidity distrubition')
```

** Most of the fixed acidity data is between 6 and 10 and the plot is right <br> skewed**
```{r echo=FALSE, message=FALSE, warning=FALSE, volatile.acidity}
ggplot(data = red, aes(x = volatile.acidity)) +
  geom_histogram() +
  ggtitle('Histogram of wine Volatile Acidity distrubition')
```

** Data points are between 0.2 and 0.8 and the plot is right skewed **

```{r echo=FALSE, message=FALSE, warning=FALSE, citric.acid}
ggplot(data = red, aes(x = citric.acid)) +
  geom_histogram() +
  ggtitle('Histogram of wine Citric Acid distrubition')
```

** Citric acid at 0.0 has the highest count next to it 0.50 **

```{r echo=FALSE, message=FALSE, warning=FALSE, residual.sugar}
ggplot(data = red, aes(x = residual.sugar)) +
  geom_histogram() +
  ggtitle('Histogram of wine Residual Sugar distrubition')
```

** Right skewed plot with most of the data points at 2 **

```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides}
ggplot(data = red, aes(x = chlorides)) +
  geom_histogram() +
  ggtitle('Histogram of wine Chlorides distrubition')
```

** Right skewed plot with most of the points arounf 0.1 **

```{r echo=FALSE, message=FALSE, warning=FALSE, free.sulfur.dioxide}
ggplot(data = red, aes(x = free.sulfur.dioxide)) +
  geom_histogram() +
  ggtitle('Histogram of wine Free Sulfur Dioxide distrubition')
```
 ** Right skewed plot with lower values having more count **
 
```{r echo=FALSE, message=FALSE, warning=FALSE, total.sulfur.dioxide}
ggplot(data = red, aes(x = total.sulfur.dioxide)) +
  geom_histogram() +
  ggtitle('Histogram of wine Total Sulfur Dioxide distrubition')
``` 

**Most of the wine samples are centered on 5, 6, 7 the extreme low and extreme <br> low can be considered as outliers**

```{r echo=FALSE, message=FALSE, warning=FALSE, fixed.acidity_plot}
ggplot(data = red, aes(x = fixed.acidity)) +
  geom_histogram() +
  ggtitle('Histogram of wine fixed acidity distrubition')
```

```{r echo=FALSE, message=FALSE, warning=FALSE,fixed.acidity_plot2}
ggplot(data = red, aes(x = fixed.acidity)) +
  geom_histogram(binwidth = .1) +
  ggtitle('Histogram of wine fixed acidity distrubition')+
  scale_x_continuous(breaks = seq(4,12,1), limits = c(4.5,13))
```


**Most of the fixed acidity data is between 6 and 10 and the plot is right <br> skewed the second plot shows the variation of the data points after using the binwidth and scale_x_continous <br>, for the second plot after zooming in using limits assuming that the points from 13 to 17 are outliers <br>, we can see fixed acidity is almost normally distribited**

```{r echo=FALSE, message=FALSE, warning=FALSE, volatile.acidity_plot}
ggplot(data = red, aes(x = volatile.acidity)) +
  geom_histogram(binwidth = 0.01) +
  ggtitle('Histogram of wine Volatile Acidity distrubition') +
  scale_x_continuous(breaks = seq(0,1.6,0.1))
```

**Data points are between 0.2 and 0.8 and the plot is right skewed**

```{r echo=FALSE, message=FALSE, warning=FALSE, citric.acid_plot}
ggplot(data = red, aes(x = citric.acid)) +
  geom_histogram(binwidth = .01) +
  ggtitle('Histogram of wine Citric Acid distrubition') +
  coord_cartesian(xlim = c(0,0.8)) +
  scale_x_continuous(breaks = seq(0,0.8,0.1))
```

**Citric acid at 0.0 has the highest count next to it 0.50, by setting the <br> binwidth to 0.01 and zooing in to take out the outliers we have a clear view over the data distribution**

```{r echo=FALSE, message=FALSE, warning=FALSE, residual.sugar_plot}
ggplot(data = red, aes(x = residual.sugar)) +
  geom_histogram(binwidth = 0.1) +
  ggtitle('Histogram of wine Residual Sugar distrubition') +
  coord_cartesian(xlim = c(0,8.1)) + 
  scale_x_continuous(breaks = seq(0,8.1,1))
```

**Right skewed plot with most of the data points at 2, I will note zoom in <br> further as I will loose a lot of data points and we have a small number of observations**

```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides_plot}
ggplot(data = red, aes(x = chlorides)) +
  geom_histogram(binwidth = 0.01) +
  ggtitle('Histogram of wine Chlorides distrubition')
```

**Right skewed plot with most of the points around 0.1**

```{r echo=FALSE, message=FALSE, warning=FALSE, free.sulfur.dioxide_plot}
ggplot(data = red, aes(x = free.sulfur.dioxide)) +
  geom_histogram(binwidth = 1) +
  ggtitle('Histogram of wine Free Sulfur Dioxide distrubition')
```

 **Right skewed plot with lower values having more count and probably any \ thing more than 40 could be an outlier**
 
```{r echo=FALSE, message=FALSE, warning=FALSE, total.sulfur.dioxide_plot}
ggplot(data = red, aes(x = total.sulfur.dioxide)) +
  geom_histogram(binwidth = 5, fill = 'orange') +
  ggtitle('Histogram of wine Total Sulfur Dioxide distrubition')
``` 

```{r echo=FALSE, message=FALSE, warning=FALSE, pH_plot}
ggplot(data = red, aes(x = pH)) +
  geom_histogram(binwidth = 0.01) +
  ggtitle('Histogram of wine pH distrubition')
``` 

**pH is nomrmally distributied**

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_plot}
ggplot(data = red, aes(x = sulphates)) +
  geom_histogram(binwidth = 0.01) +
  ggtitle('Histogram of wine Sulphate distrubition')
``` 

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_plot}
ggplot(data = red, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.1) +
  ggtitle('Histogram of wine Alcohol distrubition')
``` 



# Univariate Analysis

### What is the structure of your dataset?

12 variable (feature) with one outcome (quality)
1599 Observations
Quality is a catagorical variable ordered from 0-10 where 0 is very bad and \ 10 very excellent and needs to be converted into ordered factor.
the other 11 variables are numrics (float).

### What is/are the main feature(s) of interest in your dataset?
-Residual.sugar
-Alcohol
-volitile.acidity
-citric.acid
-quality
I will keep investigating to see a relevant other main feature in the dataset

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
-pH
-total.sulfur.dioxide
-fixed.acidity
-density

### Did you create any new variables from existing variables in the dataset?
No

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
No, data is in a well form, just a change on the data type of quality to \
become ordered factor


```{r, changing_quality_datatype}
red$quality <- as.ordered(red$quality)
```

# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, residual.sugra_alcohol}
ggplot(data = red, aes(x = residual.sugar, y= alcohol)) +
  geom_point(alpha = 1/2, position = 'jitter') +
  scale_x_continuous(breaks = seq(0,quantile(red$residual.sugar,0.99)))+
  scale_y_continuous(breaks = seq(0,quantile(red$alcohol,0.99))) +
  coord_cartesian(xlim = c(min(red$residual.sugar),
                           quantile(red$residual.sugar,0.99)),
                  ylim = c(min(red$alcohol),
                           quantile(red$alcohol,0.99)))
```

**There is no significant releation between the residual.sugra and alcohol**

```{r echo=FALSE, message=FALSE, warning=FALSE, density_alcohol}
ggplot(data = red, aes(x = density, y= alcohol)) +
  geom_point(aes(color = 'orange'), alpha = 1/4, position = 'jitter') +
  scale_x_continuous(breaks = seq(0,max(red$density),.01))+
  scale_y_continuous(breaks = seq(0,max(red$alcohol),.5)) +
  coord_cartesian(xlim = c(min(red$density),
                           quantile(red$density,0.99)),
                  ylim = c(min(red$alcohol),
                           quantile(red$alcohol,0.99))) 
```

**There is a -ve kind of strong corelationbetween alcohol and density

```{r echo=FALSE, message=FALSE, warning=FALSE, quality_volatile.acidity}
ggplot(data = red, aes(y = volatile.acidity, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .3, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$volatile.acidity,0.99),.1))+
  scale_x_continuous(breaks = seq(3,quantile(with(red,as.integer(as.vector(quality))),0.99))) +
  coord_cartesian(ylim = c(min(red$volatile.acidity),
                           quantile(red$volatile.acidity,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality))),0.99))) +
  xlab('Quality')

View(with(red,as.integer(as.vector(quality))))
```

**It seems with decrese of the volatile.acidity the quality increases on average**

```{r echo=FALSE, message=FALSE, warning=FALSE, residual.sugar_quality}
ggplot(data = red, aes(y = chlorides, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .3, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$chlorides,0.99),.1))+
  scale_x_continuous(breaks = seq(3,
                              quantile(with(red,as.integer(as.vector(quality)))
                                           ,0.99))) +
  coord_cartesian(ylim = c(min(red$chlorides),
                           quantile(red$chlorides,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality)))
                                    ,0.99))) +
  xlab('Quality')
```

**Quality at its best when the chlorides is between 0.05 and  0.1**

```{r echo=FALSE, message=FALSE, warning=FALSE, total.sulfur.dioxide_quality}
ggplot(data = red, aes(y = total.sulfur.dioxide, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .5, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$total.sulfur.dioxide,0.99),5))+
  scale_x_continuous(breaks = seq(3,
                              quantile(with(red,as.integer(as.vector(quality)))
                                           ,0.99))) +
  coord_cartesian(ylim = c(min(red$total.sulfur.dioxide),
                           quantile(red$total.sulfur.dioxide,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality)))
                                    ,0.99))) +
  xlab('Quality') +
  geom_line(fun.y = mean, stat = 'summary')
```

**lower Total sulfur dioxide (TSD) is a contributor to better quality, however \ 
I see a lower TSD with a lower  quality rating, this could be related to \ 
coorelated variable another point to be considered is that there is not enough \
samples with quality level 3, 4, 8**

```{r echo=FALSE, message=FALSE, warning=FALSE, free.sulfur.dioxide_quality}
ggplot(data = red, aes(y = free.sulfur.dioxide, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .3, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$free.sulfur.dioxide,0.99),5))+
  scale_x_continuous(breaks = seq(3,
                              quantile(with(red,as.integer(as.vector(quality)))
                                           ,0.99))) +
  coord_cartesian(ylim = c(min(red$free.sulfur.dioxide),
                           quantile(red$free.sulfur.dioxide,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality)))
                                    ,0.99))) +
  xlab('Quality')
```

**From the first look and due the not enough samples of quality level 7 and 8 \
it seems that this variable isn't contributing much to quality**

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_quality}
ggplot(data = red, aes(y = sulphates, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .3, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$sulphates,0.99),0.1))+
  scale_x_continuous(breaks = seq(3,
                              quantile(with(red,as.integer(as.vector(quality)))
                                           ,0.99))) +
  coord_cartesian(ylim = c(min(red$sulphates),
                           quantile(red$sulphates,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality)))
                                    ,0.99))) +
  xlab('Quality') +
  geom_line(stat = 'summary', fun.y = median, color = 'red') +
  geom_line(stat = 'summary', fun.y = mean, color = 'blue')
```


```{r echo=FALSE, message=FALSE, warning=FALSE, density_quality}
ggplot(data = red, aes(y = density, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .2, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$density,0.99),0.001))+
  scale_x_continuous(breaks = seq(3,
                              quantile(with(red,as.integer(as.vector(quality)))
                                           ,0.99))) +
  coord_cartesian(ylim = c(min(red$density),
                           quantile(red$density,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality)))
                                    ,0.99))) +
  xlab('Quality') +
  geom_line(stat = 'summary', fun.y = median, color = 'red') +
  geom_line(stat = 'summary', fun.y = mean, color = 'blue')
```

**This Analysis Didn't help much as I see most of the data is between 0.995 and \
0.998 of density**

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_quality}
ggplot(data = red, aes(y = alcohol, x= as.integer(as.vector(quality)))) +
  geom_point(alpha = .06, position = 'jitter') +
  scale_y_continuous(breaks = seq(0,quantile(red$alcohol,0.99),0.5))+
  scale_x_continuous(breaks = seq(3,
                              quantile(with(red,as.integer(as.vector(quality)))
                                           ,0.99))) +
  coord_cartesian(ylim = c(min(red$alcohol),
                           quantile(red$alcohol,0.99)),
                  xlim = c(3,
                           quantile(with(red,as.integer(as.vector(quality)))
                                    ,0.99))) +
  xlab('Quality')+
  geom_line(stat = 'summary', fun.y = median, color = 'red') +
  geom_line(stat = 'summary', fun.y = mean, color = 'blue')
  
```

**Alcohol does affect the quality as per the plot, as I see a trend with the \ 
increase of the alcohol**


```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_alcohol}
ggplot(data = red, aes(x = sulphates, y= alcohol)) +
  geom_point(aes(color = 'orange'), alpha = 1/5, position = 'jitter') +
  scale_x_continuous(breaks = seq(0,max(red$sulphates),.1))+
  scale_y_continuous(breaks = seq(0,max(red$alcohol),.5)) +
  coord_cartesian(xlim = c(min(red$sulphates),
                           quantile(red$sulphates,0.99)),
                  ylim = c(min(red$alcohol),
                           quantile(red$alcohol,0.99))) +
  geom_smooth()
```

**It seems that there is a relation between the sulphates and alcohol, increase \
of alcohol there is an increase of sulphates however with sulphates value of \
~0.8 the relationship changes, I assume the 0.7 and 0.8 are the optimal values**

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_free.sulfur.dioxide}
ggplot(data = red, aes(x = sulphates, y= free.sulfur.dioxide)) +
  geom_point(aes(color = 'orange'), alpha = 1/5, position = 'jitter') +
  scale_x_continuous(breaks = seq(0,max(red$sulphates),.1))+
  scale_y_continuous(breaks = seq(0,max(red$free.sulfur.dioxide),2)) +
  coord_cartesian(xlim = c(min(red$sulphates),
                           quantile(red$sulphates,0.99)),
                  ylim = c(min(red$free.sulfur.dioxide),
                           quantile(red$free.sulfur.dioxide,0.99))) +
  geom_smooth()
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

**Alcohol $ residual.sugar: even by omitting the top 1% and zooming in, I still \
 see no relation between the alcohol and residual.sugar.**
 
**Volatile.acidity & quality: I can see with the lower values of volatile.acidity \
 the quality increases**
 
**Alcohol & quality: the increase of alcohol indeed increases the quality** 

**Alcohol & density: I can see a -ve correlation between these 2 features, I will \
 these against quality to see how do they contribute to it.**
 **Quality & density: at a first look on the scatter I couldn't find a relation \
but after adding 2 geom_line of median and mean , I can tell that on average \
lower density can mean a better quality.**

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

**Chlorides & Quality: I don't see a relation between these 2.**

**Total.sulfur.dioxide & quality: generally a lower values of TSD can mean a better \
 quality this relation needs to be investigated further with more variables**
 
**Alcohol & sulphates: there is a relation especially between sulphate value \
 0.4 and 0.8**
 
**Chlorides & Quality: I don't see a relation between these 2.**
 
**Free.sulfur.dioxide & sulphates: there is no direct relation seen in the graph**

### What was the strongest relationship you found?

- Alcohol & quality
- Alcohol & density

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!