Red Wine Quality EDA by Anas Haneya
========================================================

```{r message=FALSE, warning=FALSE, include=FALSE}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(tidyr)
library(gridExtra)
library(dplyr)
library(GGally)
library(corrplot)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
#Disabling all messages, wornings and echo
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```


```{r echo=FALSE, Load_the_Data}
# Load the Data
red <- read.csv('wineQualityReds.csv')

# removing X variable as its only an index
red <- select(red, -X)
```

**Public Dataset about Red Wine with 12 variable and 1599 observation**

# Univariate Plots Section

```{r echo=FALSE, structure}
# Data structure
str(red)
```


```{r echo=FALSE, dim}
# Dataset dimensions
dim(red)
```


```{r echo=FALSE, summary}
# dataset 5 number summary
summary(red)
```

```{r echo=FALSE, nulls_check}
# nulls check
sum(is.na(red))
```

```{r function_univariate}
# Function to avoid redunduncy

uni_fun <- function(dataset, feature, title, xlabel = NULL,
                    ylabel = 'Count', binwidth = NULL) {

  return(ggplot(data = dataset, aes_string(x = feature))
              + geom_histogram(binwidth = binwidth) +
                ggtitle(title) +
                xlab(xlabel) +
                ylab(ylabel))
}
```


```{r echo=FALSE, message=FALSE, warning=FALSE, quality_plot}
# using ggplot2 library
uni_fun(red, 'quality', 'Histogram of wine qualities distrubition', 
        'Quality (0-10)', 'Count')
```

**Most of the wine samples are centered on 5, 6, 7 the extreme low and extreme \
low can be considered as outliers or not enough samples**

```{r echo=FALSE, message=FALSE, warning=FALSE, density}
#Histogram of density
uni_fun(red, 'density', 'Histogram of wine densities distrubition',
        'Density (g / cm^3)', 'Count')
```

**Density of the wine is normally distributed**

```{r echo=FALSE, message=FALSE, warning=FALSE, fixed.acidity}
#histogram of fixed.acidity
uni_fun(red, 'fixed.acidity', 'Histogram of wine fixed acidity distrubition')
```

**Most of the fixed acidity data is between 6 and 10 and the plot is right \
skewed**

```{r echo=FALSE, message=FALSE, warning=FALSE,fixed.acidity_plot2}
#fixed.acidity with a smaller binwidth and a bit of zooming
uni_fun(red, 'fixed.acidity', 'Histogram of wine fixed acidity distrubition',
        'fixed acidity (tartaric acid - g / dm^3)' , 'Count' ,0.1) +
  scale_x_continuous(breaks = seq(4,12,1), limits = c(4.5,13))
```

**Most of the fixed acidity data is between 6 and 10 and the plot is right \
skewed the second plot shows the variation of the data points after using the \ 
binwidth and scale_x_continous, for the second plot after zooming in using \
limits assuming that the points from 13 to 17 are outliers, we can see fixed \
acidity is almost normally distribited**


```{r echo=FALSE, message=FALSE, warning=FALSE, volatile.acidity}
#Histogram of volatile.acidity
uni_fun(red, 'volatile.acidity', 
        'Histogram of wine Volatile Acidity distrubition',
        'volatile acidity (acetic acid - g / dm^3)','Count')
```

**Data points are between 0.2 and 0.8 and the plot is right skewed**


```{r echo=FALSE, message=FALSE, warning=FALSE, volatile.acidity_plot}
#Histogram of volatile.acidity with a smaller binwidth
uni_fun(red, 'volatile.acidity', 
        'Histogram of wine Volatile Acidity distrubition',
        'volatile acidity (acetic acid - g / dm^3)','Count', 0.01) +
  scale_x_continuous(breaks = seq(0,1.6,0.1))
```

**Data points are between 0.2 and 0.8 and the plot is right skewed**


```{r echo=FALSE, message=FALSE, warning=FALSE, citric.acid}
#Histogram of citric acid
uni_fun(red, 'citric.acid', 
        'Histogram of wine Citric Acid distrubition',
        'citric acid (g / dm^3)','Count') 
```

**Citric acid at 0.0 has the highest count next to it 0.50**

```{r echo=FALSE, message=FALSE, warning=FALSE, citric.acid_plot}
#Histogram of citric acid with smaller binwidth and with zooming
uni_fun(red, 'citric.acid', 
        'Histogram of wine Citric Acid distrubition',
        'citric acid (g / dm^3)','Count',0.01) +
  coord_cartesian(xlim = c(0,0.8)) +
  scale_x_continuous(breaks = seq(0,0.8,0.1))
```

**Citric acid at 0.0 has the highest count next to it 0.50, by setting the \
binwidth to 0.01 and zooing in to take out the outliers we have a clear view \
over the data distribution**

```{r echo=FALSE, message=FALSE, warning=FALSE, residual.sugar}
#histogram of residual sugar
uni_fun(red, 'residual.sugar', 'Histogram of wine Residual Sugar distrubition',
        'residual sugar (g / dm^3)')
```

**Right skewed plot with most of the data points at 2**


```{r echo=FALSE, message=FALSE, warning=FALSE, residual.sugar_plot}
#histogram of residual sugar with a smaller binwidth
uni_fun(red, 'residual.sugar', 'Histogram of wine Residual Sugar distrubition',
        'residual sugar (g / dm^3)',,0.1) +
coord_cartesian(xlim = c(0.6,8.1)) + 
scale_x_continuous(breaks = seq(0.6,8.1,1))
```

**Right skewed plot with most of the data points at 2, I will note zoom in \
further as I will loose a lot of data points and we have a small number of \
observations**


```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides}
#HIstogram of Chlorides
uni_fun(red, 'chlorides', 'Histogram of wine Chlorides distrubition',
        'chlorides (sodium chloride - g / dm^3)')
```

**Right skewed plot with most of the points around 0.1**

```{r echo=FALSE, message=FALSE, warning=FALSE, free.sulfur.dioxide}
uni_fun(red, 'free.sulfur.dioxide',
        'Histogram of wine Free Sulfur Dioxide distrubition',
        'free sulfur dioxide (mg / dm^3)')
```

**Right skewed plot with lower values having more count and probably any \
 thing more than 40 could be an outlier**
 
```{r echo=FALSE, message=FALSE, warning=FALSE, total.sulfur.dioxide}
#Total Sulfur dioxide histogram with log10 transformation
uni_fun(red, 'total.sulfur.dioxide',
'Histogram of wine Total Sulfur Dioxide distrubition',
'total sulfur dioxide (mg / dm^3)') 
``` 

**Right skewed plot**

```{r echo=FALSE, message=FALSE, warning=FALSE, pH_plot}
#Histogram of the pH
uni_fun(red,'pH','Histogram of wine pH distrubition', 'pH', 'Count' ,0.01)
``` 

**pH is nomrmally distributied**

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_plot}
#Histogram of sulphates
uni_fun(red,'sulphates','Histogram of wine Sulphate distrubition', 
        'sulphates (potassium sulphate - g / dm3)','Count',0.01)
``` 

**Right skewed plot with values 0.3 to 1.0**

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_plot}
#Histogram of the alcohol percentage by volume distribution in the dataset
uni_fun(red, 'alcohol', 'Histogram of wine Alcohol distribution', 
        'alcohol (% by volume)',,0.1)
``` 


# Univariate Analysis

### What is the structure of your dataset?

- 12 variable (feature) with one outcome (quality).
- 1599 Observations.
- Quality is a catagorical variable ordered from 0-10 where 0 is very bad and \ 
10 very excellent and needs to be converted into ordered factor.
- There other 11 variables are numrics (float).

### What is/are the main feature(s) of interest in your dataset?
- Residual.sugar
- Alcohol
- Volitile.acidity
- Citric.acid
- Quality
I will keep investigating to see a relevant other main feature in the dataset

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
- pH
- total.sulfur.dioxide
- fixed.acidity
- density

### Did you create any new variables from existing variables in the dataset?
**No**

### Of the features you investigated, were there any unusual distributions? \
**Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
No, data is tidy from, and data quality is good, I just a changed data type \ 
of quality to become ordered factor and removed X variable as its not needed**


# Bivariate Plots Section

```{r, changing_quality_datatype}
#changing the quality data type to ordered factor instead of integer which fits
#the nature of this feature
red$quality <- as.ordered(red$quality)
```

```{r}
#create a function to reduce redunduncy 
#This function takes 6 parameters to dynamically plot:
# dataset (dataframe), x (x - axis), y (y-axis), xlabel, ylabel,
# alpha for scatter plot with a default value of 1
bi_fun <- function(dataset, x , y, xlabel, ylabel, alpha = 1) {
    #setting len_out to 8 to be default for non factor variables
    # len_out <- 8
    #defining df, df$c and df$y to avoid the ggplot issue if assiging the 
    #variables
    p <- NULL
    df <- dataset
    df$x = df[,which(names(df)==x)]
    df$y = df[,which(names(df)==y)]
    # converting order factor to integer for ploting
    if (class(df$x)[1] == "ordered" | class(df$x)[1] == "factor") {
        df$x <- as.integer(as.vector(df$x))
        # len_out <- 6
        p <-    ggplot(data = df, aes(x = df$x, y= df$y))+
                geom_point(alpha = alpha, position = 'jitter') +
                # scale_x_continuous(breaks = seq(min(df$x), quantile(df$x,0.99),
                #                                 length.out = len_out)) +
                # scale_y_continuous(breaks = seq(min(df$y), quantile(df$y,0.99), 
                #                                 length.out = len_out)) +
                coord_cartesian( xlim = c(min(df$x), quantile(df$x,0.99)),
                                ylim = c(min(df$y), quantile(df$y,0.99))) +
                xlab(xlabel) +
                ylab(ylabel) + 
                geom_line(stat = 'summary', fun.y = mean, aes(color = 'Mean')) +
                geom_line(stat = 'summary', fun.y = median, aes(color = 'Median')) +
                scale_color_manual(values = c('Mean' = 'red', 'Median' = 'blue')) +
                labs(color = 'Lines')
    } else if (class(df$y)[1] == "ordered" | class(df$y)[1] == "factor") {
        df$y <- as.integer(as.vector(df$y))
        # len_out <- 6
                p <-    ggplot(data = df, aes(x = df$x, y= df$y))+
                geom_point(alpha = alpha, position = 'jitter') +
                # scale_x_continuous(breaks = seq(min(df$x), quantile(df$x,0.99),
                #                                 length.out = len_out)) +
                # scale_y_continuous(breaks = seq(min(df$y), quantile(df$y,0.99), 
                #                                 length.out = len_out)) +
                coord_cartesian( xlim = c(min(df$x), quantile(df$x,0.99)),
                                ylim = c(min(df$y), quantile(df$y,0.99))) +
                xlab(xlabel) +
                ylab(ylabel) + 
                geom_line(stat = 'summary', fun.y = mean, aes(color = 'Mean')) +
                geom_line(stat = 'summary', fun.y = median, aes(color = 'Median')) +
                scale_color_manual(values = c('Mean' = 'red', 'Median' = 'blue')) +
                labs(color = 'Lines')
    } else 
       p <-    ggplot(data = df, aes(x = df$x, y= df$y))+
                geom_point(alpha = alpha, position = 'jitter') +
                # scale_x_continuous(breaks = seq(min(df$x), quantile(df$x,0.99),
                #                                 length.out = len_out)) +
                # scale_y_continuous(breaks = seq(min(df$y), quantile(df$y,0.99), 
                #                                 length.out = len_out)) +
                coord_cartesian( xlim = c(min(df$x), quantile(df$x,0.99)),
                                ylim = c(min(df$y), quantile(df$y,0.99))) +
                xlab(xlabel) +
                ylab(ylabel)
  return(p)
}
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Scatter plot of Residual.sugar & alcohol
bi_fun(red, 'residual.sugar','alcohol', 'residual sugar (g / dm^3)',
       'alcohol (% by volume)', 0.5) +
      ggtitle('Scatter plot of Residual.sugar & alcohol with a line smoother') +
      geom_smooth()
```

**There is no significant relation between the residual.sugra and alcohol**

```{r echo=FALSE, message=FALSE, warning=FALSE, density_alcohol}
# Scatter plot of Density & alcohol
bi_fun(red, 'density', 'alcohol', 'density (g / cm^3)', 'alcohol (% by volume)',
       0.25) +
  ggtitle('Scatter plot of Density & alcohol with line smoother') + 
  geom_smooth()
```

**There is a -ve kind of strong correlation between alcohol and density**

```{r echo=FALSE, message=FALSE, warning=FALSE, quality_volatile.acidity}
# Scatter plot of Quality with Volatile.acidity
bi_fun(red, 'quality','volatile.acidity', 'Quality (0-10)',
       'volatile acidity (acetic acid - g / dm^3)', 0.3) +
  ggtitle('Scatter plot of Quality with Volatile.acidity') 
```

**It seems with decrese of the volatile.acidity the quality increases on average**

```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides_quality}
# Scatter plot of Quality with chlorides
bi_fun(red, 'quality','chlorides', 'Quality (0-10)',
       'chlorides (sodium chloride - g / dm^3)', 0.3) +
  ggtitle('Scatter plot of Quality with chlorides')
```

**Quality at its best when the chlorides is between 0.05 and  0.1, also I can \
see that with the decrese of chlorides on average the quality goes up**

```{r echo=FALSE, message=FALSE, warning=FALSE, total.sulfur.dioxide_quality}
#Scatter plot of Quality with total sulfur dioxide
bi_fun(red, 'quality','total.sulfur.dioxide','quality (0-10)',
       'total sulfur dioxide (mg / dm^3)', 0.5) +
  ggtitle('Scatter plot of Quality with total sulfur dioxide')
  
```

**lower Total sulfur dioxide (TSD) is a contributor to better quality, however \ 
I see a lower TSD with a lower  quality rating, this could be related to \ 
correlated variable another point to be considered is that there is not enough \
samples with quality level 3, 4, 8**

```{r echo=FALSE, message=FALSE, warning=FALSE, free.sulfur.dioxide_quality}
# Scatter plot of free.sulfur.dioxide with quality
bi_fun(red, 'quality','free.sulfur.dioxide','quality (0-10)',
       'free sulfur dioxide (mg / dm^3)',0.3) +
  ggtitle('Scatter plot of quality with free.sulfur.dioxide')

```

**From the first look and due the not enough samples of quality level 7 and 8 \
it seems that this variable isn't contributing much to quality**

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_quality}
# Scatter plot with mean & median line of quality vs sulphates
bi_fun(red, 'quality','sulphates','quality (0-10)',
       'sulphates (potassium sulphate - g / dm3)', 0.3)
```

**Sulphates & Quality: drawing the mean and median layers showed that the \
increase of sulphates contributes to better quality.**

```{r echo=FALSE, message=FALSE, warning=FALSE, density_quality}
# Scatter plot of quality vs density with mean and median lines
bi_fun(red,'quality','density','quality (0-10)','density (g / cm^3)', 0.2)
```

**I can see that with the decrese of the density the quality improves, this is \
clear when we look at the median and mean lines**

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_quality}
# scatter plot of alcohol and quality with median and mean lines
bi_fun(red,'quality','alcohol','quality (0-10)', 'alcohol (% by volume)', 0.3)
```

**Alcohol does affect the quality as per the plot, as I see a trend with the \
increase of the alcohol**


```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_alcohol}
# scatter plot of sulphates and alcohol with smoother (liner model)
bi_fun(red,'sulphates','alcohol','sulphates (potassium sulphate - g / dm3)',
       'alcohol (% by volume)',0.2) +
  geom_smooth(method = lm)
```

**It seems that there is a relation between the sulphates and alcohol, increase\
of alcohol there is an increase of sulphates however with sulphates value of \
~0.8 the relationship changes, I assume the 0.7 and 0.8 are the optimal values**

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_free.sulfur.dioxide}
# scatter plot of sulphates and free.sulfur.dioxide with smoother
bi_fun(red,'sulphates','free.sulfur.dioxide',
       'sulphates (potassium sulphate - g / dm3)',
       'free sulfur dioxide (mg / dm^3)',0.2) +
  geom_smooth()
```

**there is no liner relation between sulphates and free.sulfur.dioxide**

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

**Alcohol $ residual.sugar: even by omitting the top 1% and zooming in, \
I still see no relation between the alcohol and residual.sugar.**
 
**Volatile.acidity & quality: I can see with the lower values of \
volatile.acidity the quality increases**
 
**Alcohol & quality: the increase of alcohol indeed increases the quality** 

**Alcohol & density: I can see a -ve correlation between these 2 features, \
I will these against quality to see how do they contribute to it.**
**Quality & density: at a first look on the scatter I couldn't find a relation \
but after adding 2 geom_line of median and mean , I can tell that on average \
lower density can mean a better quality.**

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

**Chlorides & Quality: I don't see a relation between these 2.**

**Total.sulfur.dioxide & quality: generally a lower values of TSD can mean a \
better quality this relation needs to be investigated further with more \
variables**
 
**Alcohol & sulphates: there is a relation especially between sulphate value \
 0.4 and 0.8**
 
**Chlorides & Quality: I don't see a relation between these 2.**
 
**Free.sulfur.dioxide & sulphates: there is no direct relation seen in the \
graph**

### What was the strongest relationship you found?

- Alcohol & quality
- Alcohol & density
- Suphates & quality

# Multivariate Plots Section

### Correlation

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Correlation between all features using corrplot
# converting quality to Integer to use it in corrplot package
red$quality <- as.integer(as.vector(red$quality))
co <- cor(red)
corrplot(co, type = 'full',method = 'circle')
# revert back the quality to ordered factor
red$quality <- as.ordered(red$quality)
```

**There is a significant correlation between the following: \
- citric acid & fixed acidity (Strong positive).
- citric acid & voatile acidity (Modirate negative).
- density & fixed acidity (Strong positive).
- Total sulfur dioxide & free sulfur dioxide (Strong positive).
- pH & fixed acidity (Strong negative).
- pH & citric acid (Modirate Negative).
- Density & Alcohol (Modirate Negative).
- Quality & Alcohol (Modirate Positive).**


```{r, multi_function}
#create a function to reduce redunduncy 
#This function takes 5 parameters to dynamically plot:
# dataset (dataframe), x (x - axis), y (y-axis), xlabel, ylabel

multi_fun <- function(dataset, x , y, xlabel, ylabel) {
    #setting len_out to 8 to be default for non factor variables
    # len_out <- 8
    #defining df, df$c and df$y to avoid the ggplot issue if assiging the 
    #variables
    p <- NULL
    df <- dataset
    df$x = df[,which(names(df)==x)]
    df$y = df[,which(names(df)==y)]
    # converting order factor to integer for ploting
    if (class(df$x)[1] == "ordered" | class(df$x)[1] == "factor") {
        df$x <- as.integer(as.vector(df$x))
        # len_out <- 6
        p<-ggplot(data = df, aes(x = df$x, y= df$y))+
           geom_point(alpha = alpha, position = 'jitter') +
           # scale_x_continuous(breaks = seq(min(df$x), quantile(df$x,0.99),
           #                                 length.out = len_out)) +
           # scale_y_continuous(breaks = seq(min(df$y), quantile(df$y,0.99), 
           #                                 length.out = len_out)) +
           coord_cartesian( xlim = c(min(df$x), quantile(df$x,0.99)),
                           ylim = c(min(df$y), quantile(df$y,0.99))) +
           xlab(xlabel) +
           ylab(ylabel)  
           # geom_line(stat = 'summary', fun.y = mean, aes(color = 'Mean')) +
           # geom_line(stat = 'summary', fun.y = median, aes(color = 'Median'))+
           # scale_color_manual(values = c('Mean' = 'red', 'Median' = 'blue')) +
           # labs(color = 'Lines')
    } else if (class(df$y)[1] == "ordered" | class(df$y)[1] == "factor") {
        df$y <- as.integer(as.vector(df$y))
        # len_out <- 6
        p<-ggplot(data = df, aes(x = df$x, y= df$y))+
           geom_point(alpha = alpha, position = 'jitter') +
           # scale_x_continuous(breaks = seq(min(df$x), quantile(df$x,0.99),
           #                                 length.out = len_out)) +
           # scale_y_continuous(breaks = seq(min(df$y), quantile(df$y,0.99), 
           #                                 length.out = len_out)) +
           coord_cartesian( xlim = c(min(df$x), quantile(df$x,0.99)),
                            ylim = c(min(df$y), quantile(df$y,0.99))) +
           xlab(xlabel) +
           ylab(ylabel) 
           # geom_line(stat = 'summary', fun.y = mean, aes(color = 'Mean')) +
           # geom_line(stat = 'summary', fun.y = median, aes(color = 'Median'))+
           # scale_color_manual(values = c('Mean' = 'red', 'Median' = 'blue')) +
           # labs(color = 'Lines')
    } else 
        p<-ggplot(data = df, aes(x = df$x, y= df$y))+
          # scale_x_continuous(breaks = seq(min(df$x), quantile(df$x,0.99),
          #                                 length.out = len_out)) +
          # scale_y_continuous(breaks = seq(min(df$y), quantile(df$y,0.99), 
          #                         length.out = len_out)) +
          coord_cartesian( xlim = c(min(df$x), quantile(df$x,0.99)),
                           ylim = c(min(df$y), quantile(df$y,0.99))) +
          xlab(xlabel) +
          ylab(ylabel)
  return(p)
}
```


```{r alcohol_density_quality, echo=FALSE, message=FALSE, warning=FALSE}
# a plot to explore Alcohol and density relation faceted by quality
multi_fun(red,'alcohol','density','alcohol (% by volume)','density (g / cm^3)')+
  geom_point(alpha = 0.5, position = 'jitter') +
  facet_wrap(~quality) +
  geom_smooth()
```

**I can see a the relation between the density and alcohol with quality, lower \
density higher alcohol on avergae will result a better quality**

```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# I have noticed a strong +ve correlation between fixed acidity and citric acid
# so here I am ploting these 2 against quality to see if they contribuite to it
ggplot(data = red, aes(x = fixed.acidity, y = citric.acid)) +
  geom_point(position = 'jitter', alpha = 0.2)+
  facet_wrap(~quality, scales = 'free') +
  xlab('fixed acidity (tartaric acid - g / dm^3)') +
  ylab('citric acid (g / dm^3)') +
  ggtitle('Fixed Acidity & Citric Acid faceted by Quality') +
  geom_smooth(method=lm)
```

**With the strong positive correlation beteweem fixed acidity and citric acid \
i see the quality gets better on average of the increase of both features**


```{r echo=FALSE, message=FALSE, warning=FALSE}
# I have noticed a strong +ve correlation between fixed acidity and density
# so here I am ploting these 2 against quality to see if they contribuite to it
ggplot(data = red, aes(x = fixed.acidity, y = density)) +
  geom_point(position = 'jitter', alpha = 0.2)+
  facet_wrap(~quality, scales = 'free') +
  xlab('fixed acidity (tartaric acid - g / dm^3)') +
  ylab('density (g / cm^3)') +
  ggtitle('Fixed Acidity & Density faceted by Quality') +
  geom_smooth()
```

**Eventhough there is a strong correlation between fixed acidity and density \
I see that there isn't much of relation between these 2 features and the \
quality as on average these 2 features are the same on most of quality levels**


```{r mean_groupping}
#grouping by quality and calculating the means for alcohol, 
#density,residual.sugar, volatile.acidity, citric.acid
quality_groupped <- group_by(red,quality)

groupped_quality.means <- summarise(quality_groupped, 
                                mean_alcohol = mean(alcohol),
                                mean_density = mean(density),
                                mean_residual.sugar = mean(residual.sugar),
                                mean_volatile_acidity = mean(volatile.acidity),
                                mean_citric.acid = mean(citric.acid))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, quality_all.means}
# using griupped_qiality.means to see the relationship between the main features
# and quality
ggplot(data = groupped_quality.means, aes(x = as.integer(as.vector(quality)))) +
  geom_line(aes(y = mean_alcohol ,colour = 'mean_alcohol')) +
  geom_line(aes(y = mean_residual.sugar, colour = 'mean_residual.sugar'), 
            linetype = 4) +
  geom_line(aes(y = mean_density, colour = 'mean_density'), linetype = 4) +
  geom_line(aes(y = mean_volatile_acidity, colour = 'mean_volatile_acidity'),
            linetype = 4) +
  geom_line(aes(y = mean_citric.acid, colour = 'mean_citric.acid'),
            linetype = 4) +
  scale_color_manual(values = c('mean_alcohol' = 'black', 
                                'mean_residual.sugar' = 'blue',
                                'mean_density' = 'red',
                                'mean_volatile_acidity'= 'orange',
                                'mean_citric.acid' = 'green')) +
  labs(color = 'Means') +
  xlab('Quality') +
  ylab('Means') +
  ggtitle('Means of alcohol, density, citric.acid residual.sugar and
          volatile.acidity with quality')
```

**This is a useful graph it showed me wha features affects the quality on \
average as above increasing alcohol is a strong contirbutor, also increasing \
citric acid plays a role into the quality while descreasing volatile acidity \
on the other hand, residual sugar and density alone they don't have a direct \
relation with quality**

```{r echo=FALSE, message=FALSE, warning=FALSE, cutting_pH}
## Cutting pH variable to 4 buckets to see how pH and alcohol affects quality
red$pH.bucket <- cut(red$pH, c(2.740,3.0,3.250,3.5,4.010))

# plotting Quality vs alcohol and pH as coloring
ggplot(data = red, aes(x = as.integer(as.vector(quality)), y = alcohol)) +
  geom_smooth(aes(colour = pH.bucket),se=FALSE) +
  xlab('quality (0-10)') +
  ylab('alcohol (% by volume)') +
  ggtitle('Alcohol & Quality smoother by pH.bucket')
```

**As per the above plot, I see that for pH values 3.25-3.5 and 3.5-4.01 \
and higher percentage of alcohol contirbuites to higher levels of quality**

```{r echo=FALSE, message=FALSE, warning=FALSE, cutting_tsd}
## cutting a total.sulfur.dioxide to 8 slices and plot it against alcohol,
##  desnity, total sulfur dioxide and quality
 red$tsd.bucket <- cut(red$total.sulfur.dioxide, c(0,20,40,60,80,100,150,250,289))

ggplot(data =red, aes( x = alcohol, y = density, color = tsd.bucket)) +
  geom_point(, alpha = 0.3, position = 'jitter') +
  facet_wrap(~quality) +
  xlab('alcohol (% by volume)') +
  ylab('density (g / cm^3)') +
  ggtitle('density & alcohol faceted by quality colored by total
          sulfur dioxide')
```

**following on my previous findings on total sulfur dioxide, in the above plot \
I can see buckets (20,40] and (40,60] also the relatively lower density and \
higher alcohol contributies to better quality**

```{r echo=FALSE, message=FALSE, warning=FALSE,alcohol_density_c_quality}
# convert quality to integer to have the gradient color
red$quality <- as.integer(as.vector(red$quality))

# ploting alcohol, density colored by quality
multi_fun(red,'alcohol','density','alcohol (% by volume)','density (g / cm^3)')+
  geom_point(aes(color = quality), alpha = 0.5, position = 'jitter') +
  scale_color_discrete(guide = guide_legend(reverse=TRUE)) +
  scale_color_gradient(low = 'blue', high= 'green')+ 
  ggtitle('density & alcohol colored by quality')

# revert back the quality to ordered factor
red$quality <- as.ordered(red$quality)

```

**Quality levels 6,7,8 sets on average between 11,13 of alcohol, and between \
0.996 and 0.993, so higher alcohol lower density does affect the quality of red\
wine**

```{r echo=FALSE, message=FALSE, warning=FALSE, densit_alcohol_quality_rs.bucket}

# Cutting residual.sugar to 5 buckets and plot it againts quality, alcohol and 
# density
red$residual.sugar.bucket <- cut(red$residual.sugar,c(0.9,1.9,2.2,2.6,4,15.5))

# convert quality to integer to have the gradient color
red$quality <- as.integer(as.vector(red$quality))

# plotting alcohol, density colored by quality facted by residual.sugar
multi_fun(subset(red, !is.na(residual.sugar.bucket)),'alcohol','density',
                 'alcohol (% by volume)','density (g / cm^3)') +
  geom_point(aes(color = quality) , alpha = 1, position = 'jitter') +
  scale_color_discrete(guide = guide_legend(reverse=TRUE)) + 
  facet_wrap(~residual.sugar.bucket) +
  scale_color_gradient(low = 'blue', high= 'green') +
  ggtitle('alcohol& density colored by quality facted by residual.sugar')

# revert back the quality to ordered factor
red$quality <- as.ordered(red$quality)
```

**Resudual.sugar doesn't have an influance on the quality, as in the plot**


```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_sulphates_quality}
# convert quality to integer to have the gradient color
red$quality <- as.integer(as.vector(red$quality))

# alcohol& sulphates colored by quality facted by total sulfur dioxide
multi_fun(red, 'sulphates','alcohol','sulphates (potassium sulphate - g / dm3)'
          ,'alcohol (% by volume)') +
  geom_point(aes(color = quality), alpha = .6, position = 'jitter') +
  facet_wrap(~tsd.bucket, scales = 'free') +
  scale_color_gradient(low = 'blue', high= 'green') +
  ggtitle('alcohol& sulphates colored by quality facted by total 
          sulfur dioxide')

# revert back the quality to ordered factor
red$quality <- as.ordered(red$quality)
```

**I can see in the plot alcohol with sulphates faceted total.sulfur.dioxide \
colored by quality, in addition to total.sulfur.dioxide and alcohol positive \
relation with quality increase of sulphates also contribuites to better \
quality**

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

**Adding more features againt each other allowed to explore relation between \
features of interest and againt the outcome variable, which allowed me to \
conclude what feature has the most effect and what is the optimum value for \
each**

### Interesting relation in the EDA.

**I have found out that on average a sulphates value of around 0.8 can mean a \
good quality red wine**

**Sulphates is an interesting feature I wasn't expecting to have an impact on \
red wine quality**

**while also some features are strongly correlated but it doesn't mean it will \
have a direct affect on quality**
------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
# convert quality to integer to have the gradient color
red$quality <- as.integer(as.vector(red$quality))

# alcohol& sulphates colored by quality facted by total sulfur dioxide
multi_fun(red, 'sulphates','alcohol','sulphates (potassium sulphate - g / dm3)'
          ,'alcohol (% by volume)') +
  geom_point(aes(color = quality), alpha = .6, position = 'jitter') +
  facet_wrap(~tsd.bucket, scales = 'free') +
  scale_color_gradient(low = 'blue', high= 'green') +
  ggtitle('alcohol& sulphates colored by quality facted by total 
          sulfur dioxide')

# revert back the quality to ordered factor
red$quality <- as.ordered(red$quality)
```


### Description One

**This plot shows 4 features relations and how do they contribute to the \
quality here is and explaination of the plot: \
I can see in the plot alcohol with sulphates faceted total.sulfur.dioxide \
colored by quality, in addition to total.sulfur.dioxide and alcohol positive \
relation with quality increase of sulphates also contribuites to better \
quality**

### Plot Two

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
# using griupped_qiality.means to see the relationship between the main features
# and quality
ggplot(data = groupped_quality.means, aes(x = as.integer(as.vector(quality)))) +
  geom_line(aes(y = mean_alcohol ,colour = 'mean_alcohol')) +
  geom_line(aes(y = mean_residual.sugar, colour = 'mean_residual.sugar'), 
            linetype = 4) +
  geom_line(aes(y = mean_density, colour = 'mean_density'), linetype = 4) +
  geom_line(aes(y = mean_volatile_acidity, colour = 'mean_volatile_acidity'),
            linetype = 4) +
  geom_line(aes(y = mean_citric.acid, colour = 'mean_citric.acid'),
            linetype = 4) +
  scale_color_manual(values = c('mean_alcohol' = 'black', 
                                'mean_residual.sugar' = 'blue',
                                'mean_density' = 'red',
                                'mean_volatile_acidity'= 'orange',
                                'mean_citric.acid' = 'green')) +
  labs(color = 'Means') +
  xlab('Quality') +
  ylab('Means') +
  ggtitle('Means of alcohol, density, citric.acid residual.sugar and
          volatile.acidity with quality')
```

### Description Two

**This is an excellent plot that shows on average what contributes directly to \
qualtiy which is in this case alcohol, volatile acidity, citric acid**

### Plot Three

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
# Correlation between all features using corrplot
# converting quality to Integer to use it in corrplot package
red$quality <- as.integer(as.vector(red$quality))
co <- cor(select(red, -c(pH.bucket,tsd.bucket,residual.sugar.bucket)))
corrplot(co, type = 'full',method = 'circle')
# revert back the quality to ordered factor
red$quality <- as.ordered(red$quality)
```

### Description Three

**This plot gives an idea on what to work on right away, its a starting point \
especially for the multivarite analysis, this plot show the strong, weak and \
modirate correlation between variables, as a second step in the analysis, \
I would use 2 correlated features to add one or 2 more to enchance the \
analysis**



# Reflection

**This EDA included a feature exploration of one feature at the begining \
to see how can I use them then I used bivariate analysis and ploting to further\
see the relation between most of the features in that analysis I noticed \
several feature that are correlated either positivly or negativly some of the \
features: \
- Alcohol and density.
- Quality and alcolol.
- Sulphates and quality.
- volatile acidity and quality.

This bivariate analysis allowed me to foucs on these features as starting \
point, and exploring a couple more with them.

For the multivariate anlaysis I focused on Alcohol, quality and density as \
main features then added the following features and even created some:\
- Sulphates
- pH

Newly cerated factors to help derive the relation: \
- Total sulfur dioxide bucket of 8 levels.
- pH bucket of 4 levels.
- residual sugar of 5 levels.

as conclusion I found out that pH and Total sulfur dioxide, sulphates, alcohol \
and density are the main contributors to the red wine quality.**

#####Limitiation:

**One of the struggles that I had to deal with is the dataset limitation, \
redwine quality set doesn't have a well distrubtied smaples most of the \
samples are of quality level 5 and 6, where there are no samples of level 1 \
,2, 9 and 10, and the samples on level 3, 4, 7 and 8 are very small in count \
thus, having more well balanced dataset with more samples.\

lastly, with more data, and time to explore more relation or maybe making \
more buckets can lead to more findings.\
Another thing I was thinking about is to analyse the whit wine dataset to see \
if there are any similarites between these 2 dataset, that might lead to a new \
findings.



# Refereaances: 

https://stackoverflow.com/questions/11330138/find-columns-with-all-missing-values
https://stackoverflow.com/questions/5106782/use-of-ggplot-within-another-function-in-r
